Effects of Reynolds Number on the Performance and Near-Wake Dynamics of a Vertical-Axis Cross-Flow Turbine

### 2015-04-22 paper.tex
Add comment about lack of airfoil data; resolves #12
Adjust citation of our previous work

Previously, Bachant and Wosnik \cite{Bachant2015-JoT} showed that a cross-flow
turbine wake's unique vortical mean flow field is responsible for accelerated
wake recovery when compared with conventional axial-flow propeller-type
turbines. In this study we seek to replicate the same energy balance
considerations at multiple Reynolds numbers, to examine the implications on how
small scale, i.e. low Reynolds number experiments may be used to study flows in
turbine arrays.

Bachant and Wosnik observed how maximum performance and 2-D near-wake profiles
are affected by Reynolds number \cite{Bachant2014}. In this study we set out to
analyze in greater depth the effects of scale, i.e. Reynolds number, on the
performance of the turbine, and its more detailed near-wake characteristics. We
were looking to find a threshold scale beyond which a physical model study could
reliably predict full-scale performance. We were also looking to investigate the
sensitivity as small scales, and what the implications might be for validating
the prediction capabilities of numerical models.

@@

Blade performance is generally characterized by the lift-to-drag ratio $l/d$,
which is a function of the foil's angle of attack $\alpha$. The angles of attack
for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $l/d$ is maximum---increases with blade chord Reynolds number $Re_c$. This
in general is caused by the blade boundary becoming more turbulent as a fraction
of the chord length, diminishes the negative effects of a laminar separation
bubble occurring near the leading edge. For smooth airfoils, this transition can
cause a dramatic increase in foil performance, a shown in
Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
presented in \cite{Lissaman1983}.

>>>>

Previously, is has been observed experimentally that a cross-flow turbine wake's
unique vortical mean flow field is responsible for accelerated wake recovery
when compared with conventional axial-flow propeller-type turbines
\cite{Bachant2015-JoT}. In this study we seek to replicate the same energy
balance considerations at multiple Reynolds numbers, to examine the implications
on how small scale, i.e. low Reynolds number experiments may be used to study
flows in turbine arrays.

The Reynolds number effects on the maximum performance and 2-D near-wake
profiles of the University of New Hampshire's Reference Vertical Axis Turbine
(UNH-RVAT) were presented \cite{Bachant2014}. In this study we set out to
analyze in greater depth the effects of scale, i.e. Reynolds number, on the
performance of the turbine, and its more detailed near-wake characteristics. We
were looking to find a threshold scale beyond which a physical model study could
reliably predict full-scale performance. We were also looking to investigate the
sensitivity as small scales, and what the implications might be for validating
the prediction capabilities of numerical models.

@@

Blade performance is generally characterized by the lift-to-drag ratio $l/d$,
which is a function of the foil's angle of attack $\alpha$. The angles of attack
for cross-flow turbine blades are a function of turbine tip speed ratio,
induction (slowing/turning of the free stream flow prior to reaching the
turbine), and turbine rotation or azimuthal angle $\theta$. Looking at static
airfoil data, we see that the static stall angle---the angle just beyond that at
which $l/d$ is maximum---increases with blade chord Reynolds number $Re_c$. This
in general is caused by the blade boundary becoming more turbulent as a fraction
of the chord length, diminishes the negative effects of a laminar separation
bubble occurring near the leading edge. For smooth airfoils, this transition can
cause a dramatic increase in foil performance, a shown in
Figure~\ref{fig:McMasters} at a blade chord Reynolds number on the order of
$10^5$ \cite{McMasters1980}. A review of low Reynolds number airfoil behavior is
presented in \cite{Lissaman1983}. Note that there is a distinct lack of highly
reliable data for airfoils in this transitional regime and below. An evaluation
of the various databases relevant to cross-flow turbines is presented in
\cite{Bedon2014}.
